### 2023-05-08 paper/paper.tex
Add paper structure and classification paper chapters.

% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
\pdfoutput=1
% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.

\documentclass[11pt]{article}

% Remove the "review" option to generate the final version.
\usepackage[review]{ACL2023}

% Standard package includes
\usepackage{times}
\usepackage{latexsym}

% For proper rendering and hyphenation of words containing Latin characters
% (including in bib files)
\usepackage[T1]{fontenc}
% For Vietnamese characters
% \usepackage[T5]{fontenc}
% See https://www.latex-project.org/help/documentation/encguide.pdf for other
% character sets

% This assumes your files are encoded as UTF8
\usepackage[utf8]{inputenc}

% This is not strictly necessary, and may be commented out.
% However, it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

% This is also not strictly necessary, and may be commented out.
% However, it will improve the aesthetics of text in
% the typewriter font.
\usepackage{inconsolata}
\usepackage{graphicx}

% If the title and author information does not fit in the area allocated,
% uncomment the following
%
%\setlength\titlebox{<dim>}
%
% and set <dim> to something 5cm or larger.

% Tables
\usepackage{booktabs}
\usepackage{multirow}

% Urls
\usepackage{hyperref}

% For naming embeddings
\newcommand{\Embed}[1]{\texttt{#1}}

% For naming classifiers
\newcommand{\Cls}[1]{\texttt{#1}}

% For naming transformations
\newcommand{\Trans}[1]{\textsl{`#1'}}

\title{Semantic information inside sentence embeddings}

\author{Leixin Zhang \\ 
    Computational Linguistics\\
    Seminar für Sprachwissenschaft,
    Universität Tübingen, Germany \\ 
    \texttt{leixin.zhang@student.uni-tuebingen.de}
    }
\author{David Burian\\
  Charles University, Prague\\
  \texttt{david.burian@matfyz.cz}
  }
  % Second Author \\
  % Affiliation / Address line 1 \\
  % Affiliation / Address line 2 \\
  % Affiliation / Address line 3 \\
  % \texttt{email@domain} \\}

\begin{document}
\maketitle

\begin{abstract}
This is abstract
\end{abstract}

\section{Introduction}

\section{Related work}

% Costra corpus
Costra corpus~\cite{baranvcikova2020costra}

% SBERT
SBERT~\cite{reimers2019sentence}

% Doc2vec
Doc2vec~\cite{le2014distributed}

\section{Experiments}

\subsection{Costra corpus}

% Deep dive into the properties of the datsaet
%   - transformation terminology
%   - types and explanations of transformations
%   - dataset statistics (count of seeds, sentences in total, transformation
%   distributions)

\subsection{Embeddings}\label{sec:embeddings}

% List the embeddings we are going to work with. Focus only on the core ones,
% which are used often throughout the tasks.
We focus mainly on two sets of embeddings: \Embed{doc2vec} predicted by Doc2vec
model trained on 1M of Czech sentences and \Embed{sbert} predicted by
pretrained SBERT multilingual model. Additionally we also experimented with
supervised method using the same SBERT model --- we call this embedding
\Embed{sbert\_supervised}.

% TODO: doc2vec training

To obtain \Embed{sbert\_supervised} we splitted the corpus into 5 folds. For
each fold, the model trained on the remaining 4 folds and produced embeddings
for the one remaining fold. This gave us 5 sets of disjunct embeddings, each
trained on slightly different data. For training we used softmax loss, which
classifies sentence and its corresponding seed sentence into the set of all
possible transformations. This gives the model incentive to produce embeddings
that are easily distinguishable for different transformations.

\subsection{Tasks}

% High overview of the tasks that were carried out as well as some motivation for
% doing them. Give the reader a quick overview of the work we've done.

\subsubsection{Probing tasks}

We designed a set of probing tasks which aim to test whether the embeddings
contain semantic information while ignoring how exactly it is encoded. More
specifically we ask a classifier to predict the transformation used to obtain
the given embedding.

The motivation for these tasks is that if the embedding contains semantic
information (in our case the type of transformation used), it would be
beneficial for the classifier to find and use this information. This would
boost its scores and make it standout.

We test two settings: \emph{no-context} and \emph{contextualized}. In
no-context setting the classifier receives only the sentence embedding itself
and test which (if any) transformations can be reliably determined without the
knowledge of the seed sentence. In the contextualized setting we pass both the
sentence embedding and the embedding of the corresponding seed sentence. This
enables the classifier to examine the relationships between the transformed and
seed sentences. The results are presented in Section~\ref{sec:probing_results},
the complete method is described in detail in
Appendix~\ref{appendix:probing_method}.

\section{Results}

%Enumeration of the tasks, for each task:
%  - starBriefly* describe the method used. Citing papers instead of
%  explaining.
%  - In case more information is needed (detailed method, implementation
%  notes), put it in an Appendix.
%  - Present results. Pinpoint the interesting stuff (in bold?)

\subsection{Probing for transformation prediction}\label{sec:probing_results}

Besides the already described embeddings in Section~\ref{sec:embeddings}, we
also add their versions with randomly permutated labels within sentences with
the same seed sentence (named \Embed{mixup\_by\_seed\_}) and several types
bag-of-words (\emph{BOW}) vectors. The \Embed{mixup\_by\_seed\_} embeddings
serve as a comparative random baseline. We experimented with two types of BOW
vectors --- pure bow vector having ones at positions corresponding to words
included in the given sentence (named \Embed{bow}) and TF-IDF weighted bow
vectors (named \Embed{tfidf}). Additionally we tried to limit the amount of
words registered in order to decrease the dimensionality of the vectors,
however this failed to increase the classifiers' scores and therefore we omit
them in the rest of this section. Dimensions of all embeddings are presented in
Table~\ref{tab:embeddings_dim}.

\begin{table}[htp]
  \centering
  \begin{tabular}{l  c c}
    \toprule
    Embedding & Dimension & Sparseness\\
    \midrule
    \Embed{doc2vec} & 512 & 0\%\\
    \Embed{sbert} & 384 & 0\%\\
    \Embed{bow} & 8374 & 99.89\%\\
    \Embed{tfidf} & 8374 & 99.89\%\\
    \bottomrule
  \end{tabular}
  \caption{Dimensionality and sparseness as the percentage of zeros of used
  embeddings.}\label{tab:embeddings_dim}
\end{table}

We employ several types of classifiers: Random Forests
(\Cls{RandomForestClassifier}), Support Vector Machine (\Cls{SVC}) and
K-Nearest neighbours (\Cls{KNeighboursClassifier}).

In summary, our entire pipeline is as follows:
\begin{enumerate}
  \item For each embedding, we consider only a random sample with uniform
    distribution of transformations and ignore the embeddings of the seed
    sentences.
  \item For each embedding and classifier we run a 5-fold cross-validated grid
    search over several parameters, which gives us the best model for given
    embedding.
  \item We train each embedding on the classifier which performed the best
    using cross-validation.
\end{enumerate}

More details about the pipeline can be found in
Appendix~\ref{appendix:probing_method}.

\subsubsection{Contextualized setting}

We tried two methods of contextualization: \emph{concatenating} the two
embeddings and \emph{subtracting} the seed embedding from the sentence
embedding. Since the method using difference performed better, we omit the
concatenation results.

We report the grid search results in
Figure~\ref{fig:cls_gs_diff_embed_comparison}. The best performing embeddings
were those produced by SBERT\@. Supervised SBERT embeddings differ
significantly from the unsupervised ones only in superior performance of
\Cls{KNeighboursClassifier}. This suggests better local structure in terms of
the sentence transformation.

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.5\textwidth]{figs/cls_gs_diff_embed_comparison.png}

  \caption{Grid search F1 macro averaged scores of contextualized embeddings.
  The reported scores of \Embed{sbert\_supervised} are the mean scores of all
  SBERT supervised embeddings.
  }\label{fig:cls_gs_diff_embed_comparison}

\end{figure}

Moreover the BOW embeddings surpassed \Embed{doc2vec} embedding,
proving that for prediction of transformation, lexical features are more
valuable than any information extracted by our trained Doc2vec model.

% TODO?: RandomForest performence on sparse embeddings

In Figure~\ref{fig:cls_gs_diff_labels} we show the performance of best
classifiers per each label for the most performant dense and lexical
embeddings: \Embed{sbert} and \Embed{tfidf}. We can identify four
transformations as being well-predicted having around 0.80 f1-score. However
the powerful processing of SBERT offered significant advantage to pure lexical
information only for two of those: \Trans{past} and \Trans{future}. We see even
bigger difference for \Trans{opposite meaning}, even if it reached just 0.4
f1-score.

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.5\textwidth]{./figs/cls_gs_diff_labels.png}

  \caption{Comparison of F1 score per each transformation of best classifiers
  for \Embed{sbert} and \Embed{tfidf} embedding in contextualized
  setting.}\label{fig:cls_gs_diff_labels}
\end{figure}

In order to examine the most confused transformations we summed confusion
matrices of the top 3 most performant embeddings. The resulting matrix is shown
in Figure~\ref{fig:cls_gs_diff_conf_mat}. The most confused pairs of
transformations are \Trans{minimal change} with \Trans{different meaning} and
\Trans{nonsense} with \Trans{different meaning}, both in around 40\%.
Interestingly \Trans{paraphrase} gets recognised almost never, with
\Trans{simple sentence}, \Trans{formal sentence}, \Trans{nonsense} and
\Trans{minimal change} having accuracy bellow 20\%.

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.5\textwidth]{./figs/cls_gs_diff_conf_mat.png}
  \caption{Summed confusion matrices of best classifiers for \Embed{sbert},
  \Embed{sbert\_supervised} and \Embed{tfidf} embeddings. The confusion matrix
  is row-normalized, whereas the displayed true and predicted distributions are
  not.}\label{fig:cls_gs_diff_conf_mat}
\end{figure}

\subsubsection{No-context setting}

Grid search results for no-context setting are displayed in
Figure~\ref{fig:cls_gs_no_context_embed_comparison}. Comparing these results to
those in contextualized setting, we see large drop of performance for
\Embed{sbert} embedding --- almost by 50\%. This shows how much the classifiers
relied on the information contained in the embedding of seed sentence. This
drop is not as pronounced for \Embed{sbert\_supervised} embedding, since the
model producing it is optimized for good scores in precisely this type of
task.\footnote{For this reason, we avoid comparing \Embed{sbert\_supervised} to
other embeddings from this point onwards.} On the other hand, the BOW
embeddings performed almost as well, suggesting they were not really able to
take advantage of the contextualized setting.

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.5\textwidth]{./figs/cls_gs_no_context_embed_comparison.png}

  \caption{Grid search F1 macro averaged scores of sentence embeddings with no
  context. The reported scores of \Embed{sbert\_supervised} are the mean scores
  of all SBERT supervised
  embeddings.}\label{fig:cls_gs_no_context_embed_comparison}

\end{figure}

In Figure~\ref{fig:cls_gs_no_context_labels} we compare performance of best
classifiers for \Embed{sbert} and \Embed{tfidf} embedding. Whereas the most
recognised transformations are the same as in the contextualized setting, the
classifiers performance significantly drops in all cases except for the
\Trans{ban} and the \Trans{possibility} transformations embedded with TF-IDF.
This suggests that \Trans{ban} and \Trans{possibility} transformations can be
easily identified from pure \emph{lexical} features. While SBERT fails to
encode this information into its embedding, it performs even better when the
embeddings of a sentence and its corresponding seed sentence are put into one
context. This phenomena is pronounced mainly for \Trans{different meaning} and
\Trans{past} transformations, where the inclusion of seed sentence's embedding
on the input is responsible for over 80\% and 60\% of F1 score of the
classifier, as can be seen in Figure~\ref{fig:cls_gs_labels_context_diff}.
Overall SBERT's embeddings benefit from contextualization more than TF-IDF
vectors.

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.5\textwidth]{figs/cls_gs_no_context_labels.png}

  \caption{Comparison of F1 score per each transformation of best classifiers
  for \Embed{sbert} and \Embed{tfidf} embedding in no-context
  setting.}\label{fig:cls_gs_no_context_labels}

\end{figure}

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.5\textwidth]{figs/cls_gs_labels_context_diff.png}

  \caption{Difference in F1 scores between no-context and contextualized
  setting as a percentage of contextualized
  score.}\label{fig:cls_gs_labels_context_diff}.

\end{figure}

\subsection{Evaluating Sentence Embeddings with Gaussian Mixture}

This section presents an unsupervised approach for evaluating the separability
of sentence embeddings. We measure label separability pairwisely using a
Gaussian Mixture model and calculate the F-score of the unsupervised clustered
labels with the true labels.

The Gaussian Mixture model is a probabilistic model that assumes that each
cluster follows a Gaussian (or normal) distribution and estimates the weight of
the density function for each cluster~\cite{reynolds2009gaussian,5298967}. We
assume that the sentence vectors of two distinct classes should achieve high
accuracy with Gaussian Mixture if they are displayed in a Gaussian distribution
in space and are separable from each other.

However, there is a potential limitation to using this method. The separability
and accuracy score may be underestimated if two clusters are not normally
distributed, as illustrated in Figure~\ref{fig:cirle}.

\begin{figure}[htp]
    \centering
    \includegraphics[scale=0.4]{figs/Circle.jpg}
    \caption{Cases that Gaussian Mixture Underestimate the Separability of Two
    Clusters}\label{fig:cirle}
\end{figure}

Our evaluation results show that out of 78 pairs of derivation classes, 7 pairs
achieved an accuracy score above 90\%, 24 pairs above 80\%, and 47 pairs above
60\%, as depicted in Figure~\ref{fig:GM}.

In particular, the class `ban' demonstrates good separability with many other
classes, achieving an accuracy score of 0.982 with `past', 0.980 with `formal
sentence', 0.942 with `minimal change', and 0.937 with `future', among other
pairs.

Additionally, our evaluation results reveal that tenses are generally
well-separated, with an accuracy score of 0.90 for `past' and future' classes,
and an accuracy score of 0.924 for `simple sentence' and `future' classes.

\begin{figure}[htp]
    \centering
    \includegraphics[scale=0.4]{figs/GM.png}
    \caption{Accuracy of Measured Separability with Gaussian
    Mixture}\label{fig:GM}
\end{figure}

% \subsection{SBERT as a classifier}
% This section shows whether the sentence embeddings trained on our models can
% serve as a classifier and separate sentences in terms of labels.

% For each derivation sentence, we compute the distance from its seed. The
% resulting vector. It is assumed that a certain type of derivation from a seed
% shares some similarities. A good classifier is assumed to group the same
% derivation types. For example, (sentence vector - seed vector)sentences with
% future tense are supposed to group together.
% visualize the distance vecotors

% reduced dimension PCA

% Sentence embeddings trained with SBERT can serve as classifier for some
% labels. For instance, the embeddings can separate 'future','past' and 'simple
% sentence' as in Figure X
% However, it does not work for the rest of groups:

% \subsection{Degree of Transformation}

% Whether the degree of transformation can be characterized by SBERT sentence
% embeddings. It shows the generalization group achieved most correlation. The
% more generalized sentences are further from seed sentence. While less
% generalized sentences (sentence meaning closer to seed sentences) has shorter
% distance to seed sentences

\bibliography{bibliography}
\bibliographystyle{acl_natbib}

\clearpage
\appendix
\section{SBERT embeddings}\label{appendix:sbert_embeddings}

For SBERT implementation we used the
\texttt{sentence\_transformers}\footnote{\url{https://www.sbert.net/}} python
library.

To produce sentence embeddings with SBERT, we chose pretrained model referred
to as `\textit{paraphrase-multilingual-MiniLM-L12-v2}'. It offers relatively
good scores for its size.

For the supervised SBERT sentence embeddings we list the hyperparameters used
in Table~\ref{tab:sbert_supervised_hparams}. Note that due to the small size of
training data, the model overfitted quickly. We therefore could afford to train
only for 4 epochs.

\begin{table}[htp]
  \centering
  \begin{tabular}{l c}
    \toprule
    Hyperparameter & Value\\
    \midrule
    epochs & 4 \\
    batch size & 8 \\
    warmup schedule & linear \\
    warmup steps & 10000 \\
    optimizer & AdamW \\
    weight decay & 0.1 \\
    learning rate & 2e-5 \\
    \bottomrule
  \end{tabular}

  \caption{Hyperparameters used to train SBERT for supervised
  embeddings}\label{tab:sbert_supervised_hparams}

\end{table}

\section{Probing method in detail}\label{appendix:probing_method}

For probing tasks whose results we described in
Section~\ref{sec:probing_results}, we used
\texttt{scikit-learn}\footnote{\url{https://scikit-learn.org/stable/}} python
package.

The grid searched parameters and classifiers are presented in
Table~\ref{tab:cls_gs_all_params}. We used cross-validated grid search, where
the performence of a set of parameters is evaluated using cross validation. For
all cross validations (both within and outside of grid search) we used
\texttt{StratifiedGroupKFold}, where seed identifiers were used as groups. This
caused the folds to be stratified as well as sharing as few common seed
sentences as possible.

\begin{table*}[htp]
  \centering

  \begin{tabular}{l l l}
    \toprule
    Classifier & Hyperparameter & Values searched \\
    \midrule
    \multirow{3}{13em}{\Cls{RandomForestClassifier}}
    & \texttt{n\_estimators} & 50, 100, 200 \\
    & \texttt{max\_depth} & 2, 5, 25, None \\
    & \texttt{min\_samples\_split} & 2, 10, 20 \\
    \midrule
    \multirow{2}{13em}{\Cls{SVC}*}
    & \texttt{kernel} & rbf, linear \\
    & \texttt{gamma} & auto, scale\\
    \midrule
    \multirow{2}{13em}{\Cls{KNeighboursClassifier}*}
    & \texttt{n\_neighbours} & 3, 5, 10\\
    & \texttt{weights} & uniform, distance\\
    \bottomrule
  \end{tabular}

  \caption{Classifiers and hyperparameters grid search for probe tasks. All are
  named equally as in the \texttt{scikit-learn} python package. *Embeddings
  were first scaled to have zero mean and unit
  variance.}\label{tab:cls_gs_all_params}

\end{table*}

We report the best performing set of parameters for each classifier in
Tables~\ref{tab:cls_gs_best_params_SVC},~\ref{tab:cls_gs_best_params_RandomForestClassifier}~and~\ref{tab:cls_gs_best_params_KNeighborsClassifier}

\begin{table*}[htp]
  \centering
  \input{./cls_gs_best_params_SVC_tab.tex}

  \caption{Best parameters for \Cls{SVC}
  classifier.}\label{tab:cls_gs_best_params_SVC}

\end{table*}

\begin{table*}[htp]
  \centering
  \input{./cls_gs_best_params_RandomForestClassifier_tab.tex}

  \caption{Best parameters for \Cls{RandomForestClassifier}
  classifier.}\label{tab:cls_gs_best_params_RandomForestClassifier}

\end{table*}

\begin{table*}[htp]
  \centering
  \input{./cls_gs_best_params_KNeighborsClassifier_tab.tex}

  \caption{Best parameters for \Cls{KNeighboursClassifier}
  classifier.}\label{tab:cls_gs_best_params_KNeighborsClassifier}

\end{table*}

% chosen parameters

\end{document}